### 2014-06-04 writing/S'14 paper/abstract.tex
Interim update.

\begin{abstract}
  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in a wide array of sophisticated
  language tasks, but it remains an open question whether they can
  achieve the same results as compositional semantic grammars based in
  logical forms. We address this question directly by using a logical
  grammar to generate controlled data sets encoding the relationships
  (e.g., entailment, contradiction) between pairs of expressions and
  evaluating whether each of two classes of neural model --- plain
  RNNs and recursive neural tensor networks (RNTNs) --- can learn
  those relationships correctly. Our first experiment evaluates
  whether these models can learn the basic algebra of logical
  relations involved. Our second and third experiments extend this
  evaluation to complex recursive structures and sentences involving
  quantification.  We find that the plain RNN achieves only mixed
  results on all three experiments, whereas the stronger RNTN models
  generalizes well in every setting.
\end{abstract}

>>>>

\begin{abstract}
  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in a wide array of sophisticated
  language tasks, but it remains an open question whether succeed in
  learning compositional semantic grammars that allow logical deduction. 
  We address this question directly by using a logical
  grammar to generate controlled data sets encoding the relationships
  (e.g., entailment, contradiction) between pairs of expressions and
  evaluating whether each of two classes of neural model --- plain
  RNNs and recursive neural tensor networks (RNTNs) --- can learn
  those relationships correctly. Our first experiment evaluates
  whether these models can learn the basic algebra of logical
  relations involved. Our second and third experiments extend this
  evaluation to complex recursive structures and sentences involving
  quantification.  We find that the plain RNN achieves only mixed
  results on all three experiments, whereas the stronger RNTN models
  generalizes well in every setting. This serves as compelling evidence that
  RNTNs provide a satisfactory representation of meaning for
  natural language reasoning.
\end{abstract}

### 2014-06-04 writing/S'14 paper/intro.tex
Interim update.

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in a wide array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,socher2013acl1},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging
about the ability of these models to learn compositional semantic
grammars, but it remains an open question whether they can achieve the
same results as grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013} when it comes to core semantic
concepts like quantification, entailment, and contradiction. To date,
experimental investigations of these concepts using distributed
representations have been largely confined to short phrases
\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
For robust natural language understanding, we must be able to model
these phenomena in their full generality on complex linguistic structures.

>>>>

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in a wide array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,socher2013acl1},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging
for the ability of these models to learn compositional semantic
grammars, but it remains an open question whether they can achieve the
same results as grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013} when it comes to core semantic
concepts like quantification, entailment, and contradiction. To date,
experimental investigations of these concepts using distributed
representations have been largely confined to short phrases
\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
For robust natural language understanding, we must be able to model
these phenomena in their full generality on complex linguistic structures.

### 2014-06-04 writing/S'14 paper/join.tex
First minor tweaks from the review session.
Interim update.

To test our models' ability to learn this relational structure, we
create small models for our logic in which terms denote sets of
entities from a single domain of seven entities (integers).
Figure~\ref{lattice-figure} depicts a small model of this form. The
lattice structure gives the full model, for which all the statements on
the right are valid. We divide these statements evenly into training and
test sets, and remove from the test set those statements which cannot be 
proven from the training statements using the logic depicted in 
Figure~\ref{lattice-figure}.

@@

In our experiments, we create 80 randomly generated sets drawing from
a domain of seven elements. This yields a data set consisting of
6400 statements about pairs of entities. 3200 of these pairs are
chosen as a test set, and that test set is further reduced to the 2960
examples that can be provably derived from the test data.

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models were implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer was not involved, and the two models differed only in which
layer function was used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
ensuring that the model has no information about the terms being
compared except the relations between them. We found that the RNTN
model worked best with 11-dimensional vector representations for the
80 sets and a 90-dimensional feature vector for the classifier. This
model was able to correctly label 99.3\% of the derivable test
examples, and 99.1\% of the remaining examples. The simpler RNN model
worked best with 11 and 75 dimensions, respectively, but was able to
achieve accuracies of only 90.0\% and 87.0\%, respectively.

>>>>

To test our systems's ability to learn this relational structure, we
create small models for our logic in which terms denote sets of
entities from a single domain of seven entities (integers).
Figure~\ref{lattice-figure} depicts a small model of this form. The
lattice structure gives the full model, for which all the statements on
the right are valid. We divide these statements evenly into training and
test sets, and remove from the test set those statements which cannot be 
proven from the training statements using the logic depicted in 
Figure~\ref{lattice-figure}.

@@

In our experiments, we create 80 randomly generated sets drawing from
a domain of seven elements. This yields a data set consisting of
6400 statements about pairs of entities. 3200 of these pairs are
chosen as a test set, and that test set is further reduced to the 2960
examples that can be provably derived from the training data.

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models were implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer was not involved, and the two models differed only in which
layer function was used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
ensuring that the model has no information about the terms being
compared except the relations between them.

We found that the RNTN model worked best with 11-dimensional vector representations for the
80 sets and a 90-dimensional feature vector for the classifier. This
model was able to correctly label 99.3\% of the provably derivable test
examples, and 99.1\% of the remaining test examples. The simpler RNN model
worked best with 11 and 75 dimensions, respectively, but was able to
achieve accuracies of only 90.0\% and 87.0\%, respectively.

% RNTN log: tue-j-11-6x80-hip.txt
% Train PER: 0.0021875

% RNN log: tue-j-11-6x80-r-75.txt
% Train PER: 0.047188

### 2014-06-04 writing/S'14 paper/methods.tex
Adding new Figure 1, updating results on quantifiers.
Revised figure 1
Interim update.

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the RNTN layer function
proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
nonlinearity (element-wise $\tanh$) is applied to the output of either
layer function, following \cite{socher2013acl1}.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
\end{gather} % TODO: Explain third order tensor parameter?
%
Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(l)}$ are the column vector representations
for the left and right children of the node, and $\vec{y}$ is the node's output.
The RNN concatenates them, multiplies them by an $N \times 2N$
matrix of learned weights, and applies the element-wise non-linearity
to the resulting vector (plus a bias term). The RNTN has the same
basic structure, but with the addition of a learned third-order tensor $\mathbf{T}$,
dimension $N \times N \times N$, modeling multiplicative interactions between the
child vectors. Both models include a learned bias vector, represented $\vec{b}$.

@@

\begin{figure}[htp]
  \centering
  \includegraphics[scale=0.35]{model.eps}
  \caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. 
    The same structure is used for both the RNN and RNTN layer functions.} 
  \label{sample-figure}
\end{figure}

>>>>

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the RNTN layer function
proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
nonlinearity (element-wise $\tanh$) is applied to the output of either
layer function, following \cite{socher2013acl1}.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
\end{gather} % TODO: Explain third order tensor parameter?
%
Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector representations
for the left and right children of the node, and $\vec{y}$ is the node's output.
The RNN concatenates them, multiplies them by an $N \times 2N$
matrix of learned weights, and applies the element-wise non-linearity
to the resulting vector (plus a bias term). The RNTN has the same
basic structure, but with the addition of a learned third-order tensor $\mathbf{T}$,
dimension $N \times N \times N$, modeling multiplicative interactions between the
child vectors. Both models include a learned bias vector, represented $\vec{b}$.

@@

\begin{figure}[tp]
  \centering
 % \includegraphics[scale=0.35]{model.eps}
  \input{figure1}
  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
    The same structure is used for both the RNN and RNTN layer functions.} 
  \label{sample-figure}
\end{figure}

### 2014-06-04 writing/S'14 paper/quantifiers.tex
Adding new Figure 1, updating results on quantifiers.
Interim update.

While our primary interest is in discovering the extent to which our
models can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation for the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each held out quantifier from the way that it
interacts with the other nine quantifiers which are not held
out. Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible possible
pairs of quantifiers, and we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they provide allow for the most
different class labels at in the training data. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

@@

\begin{table}[htp]
  \centering
  \begin{tabular}{ l rrr }
    \toprule
    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
    \midrule
    \textsc{all split}	& 35.4 &	60.4&	\textbf{94.4}
    \\[1ex]    
    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	\textbf{97.1} \\
    \textsc{pair not-all/not-most}	&0 &	   65.6  &	\textbf{79.2} \\
    \textsc{pair all/some}	& 0& 60.7  &	\textbf{79.5} \\
    \textsc{pair no/no}	& 30.8 &	65.8 &	\textbf{100} \\
    \bottomrule
  \end{tabular}
  \caption{Quantifier experiment performance. }
  \label{resultstable}
\end{table}

The results for these experiments are shown at the top of Figure \ref{resultstable}. We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

>>>>

While our primary interest is in discovering the extent to which our
models can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation for the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each held out quantifier from the way that it
interacts with the other nine quantifiers which are not held
out. Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible possible
pairs of quantifiers, and we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they allow for the most
different class labels at in the training data. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

@@

\begin{table}[htp]
  \centering
  \begin{tabular}{ l rrr }
    \toprule
    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
    \midrule
    \textsc{all split}	& 35.4 &	63.9&	\textbf{93.4}
    \\[1ex]    
    \textsc{pair two/less-than-two}	& 59.8 &	79.5 &	\textbf{98.3} \\
    \textsc{pair not-all/not-most}	&0 &	   62.2  &	\textbf{83.3} \\
    \textsc{pair all/some}	& 0& 63.6  &	\textbf{81.0} \\
    \textsc{pair no/no}	& 30.8 &	66.0 &	\textbf{100} \\
    \bottomrule
  \end{tabular}
  \caption{Quantifier experiment performance. }
  \label{resultstable}
\end{table}

The results for these experiments are shown in Figure \ref{resultstable}. We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

### 2014-06-04 writing/S'14 paper/recursion.tex
Interim update.

Recursive structure is a prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

>>>>

Recursive structure is a prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in training.
